---
title: "02 Make maps"
author: "DHJ"
date: "19 9 2019"
output: 
  html_document:
    keep_md: true
---


## Libs
```{r}
library(dplyr)
library(ggplot2)
library(sf)
library(readxl)

library(maps)
library(mapdata)
simple_map <- map_data("worldHires", c("Norway", "Sweden"))

library(RColorBrewer)
```

### 
```{r, fig.height = 7, fig.width = 6}
RColorBrewer::display.brewer.all()
```


## Testing ranges
```{r}
df_stations <- readxl::read_excel("01_Station_positions.xlsx") %>%
  mutate(Project = factor(Project, levels = c("Indre Oslofjord", "Ytre Oslofjord", "Økokyst"))) # for right order
         
apply(df_stations, 2, min)
apply(df_stations, 2, max)

df_stations %>% 
  ggplot(aes(LONGITUDE, LATITUDE)) +
  annotation_map(simple_map, fill = "lightgreen") +
  geom_point() +
  # geom_text(aes(label = STATION_CODE), hjust = 0, nudge_x = 0.01) +
  coord_map("lambert", parameters = c(10.6, 59.8), 
            xlim = c(9.5, 11.5), ylim = c(58.8, 59.97)) 
```

## Proper map, testing

### Test Norway (coastline) in UTM coordinates  
Problem: missing Swedish coastline....
```{r}
nc_norway <- st_read("K:/Kart/N1000/norge2.shp")
st_crs(nc_norway) <- "+proj=utm +zone=33"  # set coordinate system

gg_background <- ggplot(nc_norway) + 
  geom_sf() +
  coord_sf(xlim = c(192606, 300000), ylim = c(6548705, 6667857) - 10000, 
         expand = FALSE)
gg_background

```

### Europe map  
What is the projection?
```{r}
nc_eur <- st_read("K:/Kart/Kystshape/Europa/LinjeshapeE250")
st_crs(nc_eur) <- "+proj=longlat +ellps=WGS84"

ggplot(nc_eur) +
  geom_sf() +
  coord_sf(xlim = c(9.5, 11.5), ylim = c(58.8, 59.97), 
         expand = FALSE)

head(nc_eur)

x <- sf::st_sample(nc_eur, 1000)
str(x, 1)

```

## Proper map, chosen map

### Coastline data from OpenStreetmap  
Stored as a simple data frame of x and y coordinates, with NAs separating features (e.g. islands)
* So we skip using sf at all, we just plot points as usual in ggplot  
* Coastline is added using geom_path(); line is broken up at NAs   
* We pre-filter the data to increase speed of plotting, but don't throw away the NAs
* Can still get correct projected drawing, using coord_map()   
```{r}
fn <- "K:/Avdeling/214-Oseanografi/DHJ/Data/Maps/coastline_skagerrak_coords_longlat.RData"
load(fn)  # loads coastline_skagerrak_coords_longlat

xlim <- c(9.5, 11.5)
ylim = c(58.8, 59.97)

gg <- 
  coastline_skagerrak_coords_longlat %>%  
  filter(is.na(x) | 
           (x > xlim[1] & x < xlim[2] & y > ylim[1] & y < ylim[2])) %>%   # prefiltering speeds up drawing
  ggplot(aes(x,y)) +
  geom_path(size = 0.35, na.rm = TRUE) + 
  coord_map("lambert", parameters = c(10.6, 59.8), 
            xlim = xlim, ylim = ylim) +
  theme()

ggsave("Figures/02_test04.png", gg)

```

## Plotting

### Prepare  
Colors + coastline data
```{r}
colors <- RColorBrewer::brewer.pal(8, "Set1")[c(1,2,5)]

df_map <- coastline_skagerrak_coords_longlat %>%
   # prefiltering speeds up drawing, but remembr to NOT throw NAs away! (Ssee '02_test01.png')
   filter(is.na(x) | (x > xlim[1] & x < xlim[2] & y > ylim[1] & y < ylim[2]))    drawing
```


### Version 1 plots  
With points without borders  
Map 1,2,3 are for 1) Indre Oslofjord, 2) Indre + Ytre Oslofjord, 3) all three programs  
20 seconds per plot  
```{r}

gg <- df_stations %>% 
  filter(Project %in% "Indre Oslofjord") %>%
  ggplot(aes(LONGITUDE, LATITUDE)) +
  geom_path(data = df_map, aes(x,y), color = "grey40", size = 0.3, na.rm = TRUE) + 
  geom_point(aes(color = Project), size = 2.5) +
  scale_color_manual(values = colors) +
  coord_map("lambert", parameters = c(10.6, 59.8), 
            xlim = c(9.5, 11.5), ylim = c(58.8, 59.97)) +
  theme(axis.title = element_blank())
# gg

ggsave("Figures/02_map1_version1.png", gg)


gg <- df_stations %>% 
  filter(Project %in% c("Indre Oslofjord", "Ytre Oslofjord")) %>%
  ggplot(aes(LONGITUDE, LATITUDE)) +
  geom_path(data = df_map, aes(x,y), color = "grey40", size = 0.3, na.rm = TRUE) + 
  geom_point(aes(color = Project), size = 2.5) +
  scale_color_manual(values = colors) +
  coord_map("lambert", parameters = c(10.6, 59.8), 
            xlim = c(9.5, 11.5), ylim = c(58.8, 59.97)) +
  theme(axis.title = element_blank())
# gg

ggsave("Figures/02_map2_version1.png", gg)


gg <- df_stations %>% 
  ggplot(aes(LONGITUDE, LATITUDE)) +
  geom_path(data = df_map, aes(x,y), color = "grey40", size = 0.3, na.rm = TRUE) + 
  geom_point(aes(color = Project), size = 2.5) +
  scale_color_manual(values = colors) +
  coord_map("lambert", parameters = c(10.6, 59.8), 
            xlim = c(9.5, 11.5), ylim = c(58.8, 59.97)) +
  theme(axis.title = element_blank())
# gg

ggsave("Figures/02_map3_version1.png", gg)

```


### Version 2 plots   
With points with black borders (pch = 21, and using 'fill' instead of 'color')  
```{r}

gg <- df_stations %>% 
  filter(Project %in% "Indre Oslofjord") %>%
  ggplot(aes(LONGITUDE, LATITUDE)) +
  geom_path(data = df_map, aes(x,y), color = "grey40", size = 0.3, na.rm = TRUE) + 
  geom_point(aes(fill = Project), pch = 21, size = 2.5) +
  scale_fill_manual(values = colors) +
  coord_map("lambert", parameters = c(10.6, 59.8), 
            xlim = c(9.5, 11.5), ylim = c(58.8, 59.97)) +
  theme(axis.title = element_blank())
# gg

ggsave("Figures/02_map1_version2.png", gg)


gg <- df_stations %>% 
  filter(Project %in% c("Indre Oslofjord", "Ytre Oslofjord")) %>%
  ggplot(aes(LONGITUDE, LATITUDE)) +
  geom_path(data = df_map, aes(x,y), color = "grey40", size = 0.3, na.rm = TRUE) + 
  geom_point(aes(fill = Project), pch = 21, size = 2.5) +
  scale_color_manual(values = colors) +
  coord_map("lambert", parameters = c(10.6, 59.8), 
            xlim = c(9.5, 11.5), ylim = c(58.8, 59.97)) +
  theme(axis.title = element_blank())
# gg

ggsave("Figures/02_map2_version2.png", gg)


gg <- df_stations %>% 
  ggplot(aes(LONGITUDE, LATITUDE)) +
  geom_path(data = df_map, aes(x,y), color = "grey40", size = 0.3, na.rm = TRUE) + 
  geom_point(aes(fill = Project), pch = 21, size = 2.5) +
  scale_fill_manual(values = colors) +
  coord_map("lambert", parameters = c(10.6, 59.8), 
            xlim = c(9.5, 11.5), ylim = c(58.8, 59.97)) +
  theme(axis.title = element_blank())
# gg

ggsave("Figures/02_map3_version2.png", gg)

```

